\tcbsubtitle{Simple}

\tcbsubtitle{High Availability}

\tcbsubtitle{General}

1. Etcd Backup \& Restore 💾

The etcd database holds the entire state of your cluster. You must know how to back it up using etcdctl.

    Snapshot save: ```bash
    \begin{verbatim}
        ETCDCTL_API=3 etcdctl --endpoints=https://www.google.com/search?q=https://127.0.0.1:2379
    \end{verbatim}

    --cacert=<ca-file> --cert=<cert-file> --key=<key-file>

    snapshot save /opt/snapshot-pre-boot.db


2. Upgrading the Cluster ⬆️
    The exam often asks you to upgrade a node (e.g., from v1.29 to v1.30). The sequence is critical:

    Drain the node (move Pods off safely).

    Upgrade the kubeadm tool.

    Upgrade the node configuration (kubeadm upgrade apply or node).

    Upgrade kubelet and kubectl.

    Uncordon the node (let Pods back in).

A Troubleshooting Scenario:
You are asked to check why a node is showing a status of NotReady. You SSH into the node and notice that the Kubelet service is stopped.


On the CKA exam, the standard command you’ll want to have in your muscle memory is:
Bash

systemctl restart kubelet

And if you need to make sure it starts automatically when the node boots up:
Bash

systemctl enable kubelet

For this domain of your cheatsheet, remember that the Kubelet is the only component that runs as a regular system service. The others (API Server, Scheduler, etc.) usually run as Static Pods.

If those other components are failing, you won't find them in systemctl; instead, you'll need to look at the manifest files located in /etc/kubernetes/manifests/.


3. Upgrades (kubeadm)

    apt upgrade kubeadm

    kubeadm upgrade plan

    kubeadm upgrade apply vX.Y.Z

    apt upgrade kubelet kubectl \& systemctl restart kubelet


