Networking	Services, Ingress vs Gateway API, DNS, and Network Policies.

The Three Main Service Types
Type	Use Case	Reachability
ClusterIP	Internal communication (default).	Only inside the cluster.
NodePort	Exposing a service on a static port on each Node's IP.	Outside the cluster (via NodeIP:Port).
LoadBalancer	Using a cloud provider's external load balancer.	Outside the cluster (via Public IP).
Quick Networking Commands

    Expose a deployment: kubectl expose deployment nginx --port=80 --target-port=80 --type=ClusterIP

    Check endpoints: kubectl get endpoints (This shows you which Pod IPs are currently "behind" your service).

Network Policies 🛡️
the moment you apply a NetworkPolicy to a Pod, that Pod becomes "isolated."

By default, Kubernetes uses an "Allow All" posture. However, as soon as a selector matches a Pod in a NetworkPolicy, that Pod switches to a "Default Deny" posture for the types of traffic (Ingress or Egress) defined in the policy.

🛠️ Key NetworkPolicy Commands

Since you can't create these easily with kubectl create, you'll usually copy a template from the official docs during the exam.

    Check existing policies: kubectl get netpol

    Describe a policy: kubectl describe netpol <name> (This is vital to see which Pods are actually being targeted by the selector).



🌐 Advanced Networking: Ingress

While we talked about Services (ClusterIP/NodePort), Ingress is how you manage external access to those services, typically via HTTP/HTTPS. It acts as a smart router.

An Ingress resource needs two things to work:

    An Ingress Controller (the software, like NGINX, that does the work).

    An Ingress Resource (the "rules" you write).


Imagine you have two services: blue-service and green-service. You want traffic going to myapp.com/blue to hit the blue service, and myapp.com/green to hit the green service.

In the Ingress YAML, what is the name of the field where you define these URL patterns (like /blue or /green)?

It’s part of a hierarchy that looks like this: rules -> http -> paths -> path.

When you define a path, you also have to specify a pathType. For the CKA, you'll mostly use Prefix (matches based on a URL prefix) or Exact.

Here is a quick look at how that structure comes together for your "Blue/Green" scenario:
YAML

spec:
  rules:
  - host: myapp.com
    http:
      paths:
      - path: /blue
        pathType: Prefix
        backend:
          service:
            name: blue-service
            port:
              number: 80
      - path: /green
        pathType: Prefix
        backend:
          service:
            name: green-service
            port:
              number: 80



In Kubernetes, every Service gets a DNS name that other Pods can use to find it. This prevents you from needing to hardcode IP addresses. The format is usually:
service-name.namespace.svc.cluster.local




    ClusterIP: Internal only.

    NodePort: Exposed on 30000-32767.

    Ingress: HTTP/HTTPS routing. Remember the pathType: Prefix and backend structure.