Storage	PV/PVC Lifecycle, AccessModes, StorageClasses, and CSI.


💾 The Storage Lifecycle

To remember the flow, think of it as a three-step process:

    PersistentVolume (PV): The actual "chunk" of storage (like a physical disk or NFS share) created by the cluster admin. It exists at the cluster level.

    PersistentVolumeClaim (PVC): A request for storage by a user. It lives in a specific Namespace.

    Pod Volume: The Pod "mounts" the PVC to a specific directory inside the container.

🛠️ Quick Storage Commands

    Check storage classes: kubectl get sc (Use this to see if "dynamic provisioning" is available).

    List all storage components: kubectl get pv,pvc


Storage is often the most "manual" part of the CKA. You have to ensure the Access Modes and Capacity match between the provider and the user.
Component	Responsibility	Key Field to Watch
PersistentVolume (PV)	Cluster Admin	capacity, accessModes, hostPath or nfs
PersistentVolumeClaim (PVC)	Developer/App	resources.requests.storage, accessModes
Pod	Workload	volumes.persistentVolumeClaim.claimName

A Common Pitfall:
If you create a PVC requesting 10Gi of storage, but your only available PV is 5Gi, what do you think the status of that PVC will be when you run kubectl get pvc?
The PVC will stay in a Pending state indefinitely because the "contract" between the claim and the volume cannot be fulfilled.

In the CKA, this is a common troubleshooting task. If you see a PVC stuck in Pending, you should immediately check:

    Capacity: Does the PV have enough space?

    AccessModes: Do they match (e.g., ReadWriteOnce vs ReadOnlyMany)?

    StorageClass: Are they using the same class name?

